---
title: "DevDataProductsPA1"
author: "Jeff Sternin"
date: "February 25, 2018"
output: html_document
---

## Path from east to west between Top 10 US Cities  

The goal to show in the map of USA Top 10 cities (by population) and connect them with the path from
east (New York) to west (San Jose), show coordinates, population and the distance to next city on popup.
See result here <https://jsternin.github.io/DevDataProdPA1/>


```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
library(leaflet)
library(ggmap)
library(maps)
```

```{r calculate_distance,echo=TRUE,include=TRUE}
getDistanceFromLatLng <- function (lat1, lng1, lat2, lng2) { 
  ## modified for R from here https://simplemaps.com/resources/location-distance
  deg2rad <- function (deg){deg * (pi/180);}
  square <- function (x){x^2;}
  r <- 6371; # radius of the earth in km
  lat1=deg2rad(lat1);
  lat2=deg2rad(lat2);
  lat_dif<-lat2-lat1;
  lng_dif<-deg2rad(lng2-lng1);
  a<-square(sin(lat_dif/2))+cos(lat1)*cos(lat2)*square(sin(lng_dif/2));
  d<-2*r*asin(sqrt(a));
  d * 0.621371;
}
```

Load data (coordinates, population, path)

```{r get_data,echo=TRUE,include=TRUE}

citinames <- c("New York,New York","Chicago,Illinois","Los Angeles, California",
               "Houston, Texas","Philadelphia, Pennsylvania", "Phoenix, Arizona","San Antonio, Texas",
               "San Diego, California","Dallas, Texas","San Jose, California")
coord = data.frame(lon = c(-73.968285,-87.62980,-118.24368, -95.36980,-75.16522,
                           -112.07404,-98.49363, -117.16108, -96.79699,-121.88633), #-115.17456
                   lat = c(40.785091, 41.87811, 34.05223, 29.76043, 39.95258, 
                               33.44838, 29.42412, 32.71574, 32.77666, 37.33821)) #36.10237
path =     c(1,5,2,9,4,7,6,8,3,10) ## east-west path 
next2west = c(1:10) ## wext city to the west from ith ranked city
for(i in 1:length(path)) { 
  if (path[i] < length(path)) 
    next2west[path[i]] = path[i+1]
  else
    next2west[path[i]] = path[i]
}
##next2west =c(5,9,10,7,2,8,6,3,4,10) ## next westward city on the path 
path_lon<-NULL;path_lat<-NULL;dist2west<-NULL;
for(i in 1:10) dist2west <- c("",dist2west)
for(i in 1:length(path)) {
  i1 <- path[i]; 
  path_lon <- c(path_lon,coord[i1,1]);
  path_lat <- c(path_lat,coord[i1,2])
  i2 <- next2west[i]
  dist <- getDistanceFromLatLng(coord[i,2],coord[i,1],coord[i2,2],coord[i2,1])
  str <- sprintf("Dist.to %s: %5.2f miles", citinames[i2],dist)
  dist2west[i] <- str
}
east_west<- data.frame(lon=path_lon,lat=path_lat,dist2west=dist2west)
top10cities <- data.frame( name = citinames,
    population = c(8405837,3884307,2718782,2195914,1553165,1513367,1409019,1355896,1257676,998537),
    Coord = coord ) 

print("Top 10 US Cities by population")
print(top10cities)
```


```{r get_coord,echo=TRUE,include=FALSE}
##top10citiesCoord<-geocode(citinames)  ## get lat/long coordinates.
```

Find center of the map and zoom. Add Circle markers with radius proportional to city population and other
info about city.

```{r do_map,echo=TRUE,include=TRUE,fig.width=10,fig.height=5}
cntlon = (min(coord$lon,na.rm=TRUE)  + max(coord$lon,na.rm=TRUE))/2
cntlat = (min(coord$lat,na.rm=TRUE) + max(coord$lat,na.rm=TRUE))/2
coord %>% leaflet() %>% setView(cntlon, cntlat, zoom = 4) %>% addTiles() %>%
addCircleMarkers(popup=sprintf(
  "Population: %d, Latitd: %5.2f, Lontitd: %5.2f %s",
   top10cities$pop,top10cities$Coord.lat,top10cities$Coord.lon,east_west$dist2west),
    label=top10cities$name, radius = sqrt(top10cities$pop/2000),
  clusterOptions=markerClusterOptions()) %>%  
  addPolylines(lat = east_west$lat, lng = east_west$lon)

```


